---
title: "Nested Dirichlet Analysis of Social Vulnerability Index"
author: "Erika Shults and Monnie McGee"
format: pdf
editor: visual
embed-resources: true
---

## Texas Counties SVI Analysis

Explain the SVI. What are the variables and how are they measured (give units), how is it used, etc.

Literature Review.

Why are we doing this? What insights do we gain?

## Set Up the data

R packages necessary

```{r}
#| label: R packages
#| echo: false
library(SVI)
library(compositions)
library(coda.base)
library(DirichletReg)
library(sqldf)
library(GGally)
library(ggtern)
```

The file "Texas_county.csv" should be in the current working directory. What is in the file Texas_county.csv?

```{r}
##Read in the data##
texas <- read.csv("Texas_county.csv")
head(texas)
```

There are 254 counties in Texas and we have information on `r nrow(texas)` of them.

The following code uses SQL to create a clean data set with only the county and corresponding ranking information. Ranking information for what?

```{r}
tx <- sqldf('select 
              STCNTY,
              COUNTY as county, 
              Rep_Dem,
              EPL_POV150,
              EPL_UNEMP,
              EPL_HBURD,
              EPL_NOHSDP,
              EPL_UNINSUR,
              EPL_AGE65,
              EPL_AGE17,
              EPL_DISABL,
              EPL_SNGPNT,
              EPL_LIMENG,
              EPL_MINRTY,
              EPL_MUNIT,
              EPL_MOBILE,
              EPL_CROWD,
              EPL_NOVEH,
              EPL_GROUPQ,
              SPL_THEME1,
              SPL_THEME2,
              SPL_THEME3,
              SPL_THEME4,
              SPL_THEMES,
              RPL_THEME1,
              RPL_THEME2,
              RPL_THEME3,
              RPL_THEME4,
              RPL_THEMES
      from texas')
head(tx)
```

We have information on `r nrow(tx)` counties.

Next, we examine whether there is a difference in overall SPL_THEMES between Reps and Dems.

Are Reps and Dems at the county level (how are Reps and Dems defined?) What are SPL_THEMES?

```{r}
aggregate(x = tx$SPL_THEMES , by = list(tx$Rep_Dem), FUN = mean) 
ggplot(tx, aes(x=Rep_Dem, y=SPL_THEMES)) + geom_boxplot()
```

From the ANOVA, it looks like there is a difference overall. Democratic counties tend to have higher overall SVI (higher indicates more vulnerable).

```{r}
res_aov <- aov(SPL_THEMES ~ Rep_Dem, data = tx)
summary(res_aov)
```

These are residual plots for the ANOVA to check assumptions.

```{r}
plot(res_aov)
```
